Question 2
======================================================================

__Goal__: infer the average activity pattern over time (within a day)

Written by [Russo](https://github.com/russojhsph/fitbitR).

To make it easier for me to write this document, the interpretation of each plot is included as comments in the R code. So toggle the R code to view the comments and interpretation if you are curious.

# Setup

```{r setup}
## Packages used
library(fitbitR)
require(ggplot2)
require(mgcv)

## Load the data
data <- preprocess(fitbitData)

## Complete predictions
types <- c("lm", "poisson", "means", "overall-mean")
datap <- lapply(types, function(x) { fitbitPred(data, method=x)})

## All data
all <- c(list(data), datap)
names(all) <- c("original", types)
```

# Mean by interval

```{r meanint}

## Mean by interval
intMean <- function(data) {
	tapply(data$nSteps, data$Interval, mean, na.rm=TRUE)
}
meanint <- lapply(all, intMean)

## Plot the mean for each data set
lapply(names(meanint), function(x) { plot(meanint[[x]], main=x, xlab="IntS", ylab="Number of steps", type="o")})
```

This gives us an idea of the overall activity pattern by day. 


# GAM

Another approach is to use a GAM model which can lead into question 3 where we will want to check if there is a difference between weekends and weekdays. The GAM model will smooth the data and will have some trouble fitting the high-peaks, but will overall give a good idea of the average activity pattern by day.

```{r gamind}

## Transform the Interval to a scale in 0 to 287: added this to preprocess()
## Initial GAM using the poisson family.
ggplot(data, aes(x=IntS, y=nSteps)) + geom_point(size=0.5, na.rm=TRUE) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=poisson)

## Again but with a quasipoisson family which should be better given the overdispersion detected earlier.
ggplot(data, aes(x=IntS, y=nSteps)) + geom_point(size=0.5, na.rm=TRUE) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)

## The same but using Interval instead of IntS. The curves look similar but it's easier to visualize the actual time of the day.
ggplot(data, aes(x=Interval, y=nSteps)) + geom_point(size=0.5, na.rm=TRUE) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)
```

In summary, this person has a high peak of activity between 6 and 9 am. Then it stabilizes from 10 am to 4 pm. There is a slight increase at 5 and 6 pm and decreases until very low levels at 10 pm. The person is rarely active from 11 pm to 5 am.


# Using q2()

Now I can use q2() to get the activity pattern either using the mean or the gam method.

```{r q2fun}
q2.mean <- q2(data, "mean")
q2.gam <- q2(data, "gam")
```

The mean method is much more variable than the GAM method and it is harder to interpret. However, the large activity peak is more noticable.



# Reproducibility

```{r reproducibility}
sessionInfo()
print(proc.time())
```


