Exploratory data analysis
=========================

Written by [Russo](https://github.com/russojhsph/fitbitR).

To make it easier for me to write this document, the interpretation of each plot is included as comments in the R code. So toggle the R code to view the comments and interpretation if you are curious.

# Setup

```{r setup}
## Packages used
library(fitbitR)
require(ggplot2)
```

# Basic exploration

```{r basicExploration}
## Load the data
data <- preprocess(fitbitData)

## Note that the number of rows corresponds to ncol * nrow from fitbitData
dim(data); ncol(fitbitData) * nrow(fitbitData)
head(data)

## Note that there are 2304 NA's because 8 days have missing data.
summary(data)
sum(is.na(data$nSteps)) / nrow(fitbitData)
```

# Number of steps distribution

```{r nSteps}
## Histogram and density plot for the number of steps at a given interval
hist(data$nSteps, freq=FALSE, col="light blue")
lines(density(data$nSteps[!is.na(data$nSteps)]), col="red")

## Histogram with log-transformed data. There seem to be two modes.
hist(log(data$nSteps), freq=FALSE, col="light blue")
lines(density(log(data$nSteps[!is.na(data$nSteps)])), col="red")

## Could the number of steps look like a Poisson?
## Estimate lambda
l <- mean(data$nSteps, na.rm=TRUE)

## Does the mean look very different from the variance? Yes, it's very different.
l
var(data$nSteps, na.rm=TRUE)

## Generate some random poisson data to compare against the observed data
set.seed(20130605)
rand <- rpois(nrow(data), l)

## Doesn't look like poisson dist is kept at all.
qqplot(rand, data$nSteps)
qqline(data$nSteps, distribution = function(p) qpois(p, lambda = l) )

## Normal looking in log scale? Nope.
qqnorm(log(data$nSteps + 1))

```


# Displaying with points

```{r scatter}

## Ok as an entry plot. Can clearly notice the NA's.
ggplot(data, aes(y=Date, x=Interval, colour=nSteps)) + geom_point()

## The one issue here is that NA's look like high activity days
ggplot(data, aes(y=Date, x=Interval, colour=Weekend, alpha=nSteps)) + geom_point()

## Size can get a tad big and overlap other days
ggplot(data, aes(y=Date, x=Interval, colour=Weekend, size=nSteps)) + geom_point()

## I like this plot. It's easy to notice the NA's, the weekday vs weekend and the activity peak: 8 am on weekdays.  Specially for 2 and 3rd week of October.
ggplot(data, aes(y=Date, x=Interval, colour=nSteps)) + geom_point() + facet_grid(~ Weekend)

## NA's are not visible, but the activity peaks are easier to note. 
ggplot(data, aes(x=Interval, y=nSteps, group=Date, colour=Day, alpha=nSteps)) + geom_point() + facet_grid(~Weekend)
```

# Displaying with lines

```{r curves}

## This one summarizes all the information across different dates. Colours overlap too much.
ggplot(data, aes(x=Interval, y=nSteps, colour=Weekend)) + geom_line()

## An improvement, now summarizing by day of the week. It's easy to see the winner by interval, but not the information within it due to colours overlapping.
ggplot(data, aes(x=Interval, y=nSteps, colour=Day)) + geom_line(alpha=1/2) + facet_grid(~Weekend)

## A rather unuseful plot.
ggplot(data, aes(x=Interval, y=nSteps)) + geom_line() + facet_grid(~Weekend)

## Colour helps, but the information looked better when using points.
ggplot(data, aes(x=Interval, y=nSteps, group=Date, colour=Day)) + geom_line() + facet_grid(~Weekend)

## It looks a tad better, but you can't really notice much the difference by dates.
ggplot(data, aes(x=Interval, y=nSteps, group=Date, colour=Weekend)) + geom_line() + facet_grid(~ Day)
```

# Polar display

```{r polar}
## This plot is good for noticing the difference by hour (interval) and the overall maximum by merging across dates.
ggplot(data, aes(x=Interval, y=nSteps, colour=Weekend)) + geom_line() + coord_polar()

## Messier plot and it's harder to notice the differences though the variability is more apparent.
ggplot(data, aes(x=Interval, y=nSteps, colour=Weekend, group=Date)) + geom_line() + coord_polar()

## I think that this is one of the plots that I like the most so far. It further improves the points plot from earlier by taking advantage of the polar display to clearly notice the hours of no activity..
ggplot(data, aes(x=Interval, y=nSteps, group=Date, colour=Day, alpha=nSteps)) + geom_point()  + coord_polar() + facet_grid(~Weekend)
```



# Reproducibility

```{r reproducibility}
sessionInfo()
print(proc.time())
```


