Question 3
====================================================================================================

__Goal__: infer whether there are difference between weekdays and weekends in patterns of activity

Written by [Russo](https://github.com/russojhsph/fitbitR).

To make it easier for me to write this document, the interpretation of each plot is included as comments in the R code. So toggle the R code to view the comments and interpretation if you are curious.

# Setup

```{r setup}
## Packages used
library(fitbitR)
require(ggplot2)
require(mgcv)

## Load the data
data <- preprocess(fitbitData)

## Complete predictions
types <- c("lm", "poisson", "means", "overall-mean")
datap <- lapply(types, function(x) { fitbitPred(data, method=x)})

## All data
all <- c(list(data), datap)
names(all) <- c("original", types)
```

# GAM

This section is inspired by Q2.

## Separating by weekday vs weekend

```{r gamweekend, fig.width=10}
## Fitting a GAM (poisson) by interval with data grouped by Day.
ggplot(data, aes(x=Interval, y=nSteps, group=Weekend)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Weekend) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=poisson)
ggplot(data, aes(x=IntS, y=nSteps, group=Weekend)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Weekend) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=poisson)

## Same but with quasipoisson.
ggplot(data, aes(x=Interval, y=nSteps, group=Weekend)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Weekend) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)
ggplot(data, aes(x=IntS, y=nSteps, group=Weekend)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Weekend) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)
ggplot(data, aes(x=IntS, y=nSteps, group=Weekend)) + geom_point(alpha=1/4, na.rm=TRUE) + facet_grid(~Weekend) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)
ggplot(data, aes(x=IntS, y=nSteps, group=Weekend)) + geom_point(alpha=1/4, na.rm=TRUE, colour="orange") + facet_grid(~Weekend) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)
```

## By interval & day of the week


```{r gamdow, fig.width=12}
## Fitting a GAM (poisson) by interval with data grouped by Day.
ggplot(data, aes(x=Interval, y=nSteps, group=Day)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Day) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=poisson)
ggplot(data, aes(x=IntS, y=nSteps, group=Day)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Day) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=poisson)

## Same but with quasipoisson.
ggplot(data, aes(x=Interval, y=nSteps, group=Day)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Day) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)
ggplot(data, aes(x=IntS, y=nSteps, group=Day)) + geom_point(size=0.5, na.rm=TRUE) + facet_grid(~Day) + stat_smooth(method=gam, formula= y ~ s(x, bs="cr"), na.rm=TRUE, family=quasipoisson)
```

## Model

Next I fit a GAM model with the poisson and quasipoisson families adding the Weekend as a covariate.

```{r model}
fit <- gam(nSteps ~ s(IntS, bs="cr") + Weekend, data=data, family=poisson)
summary(fit)
fit2 <- gam(nSteps ~ s(IntS, bs="cr") + Weekend, data=data, family=quasipoisson)
summary(fit2)
```

* The quasipoisson seems like a better option because it estimates the variance which is a problem as seen before due to the overdispersion.
* There is a significant difference between Weekdays and Weekends.

```{r q3ci}
## CI for the Weekend coef
est <- fit2$coef["WeekendWeekend"] 
est.sd <- sqrt(vcov(fit2)["WeekendWeekend","WeekendWeekend"])
t <- qt(0.975, df=fit2$df.null)
## I'm getting the CI without /sqrt(n) because it's the SE already.
exp(est + c(-1, 1) * t * est.sd)
```


## Prediction

Next I check whether the results are influenced by using the data with predictions.

```{r modelpred}

fits <- lapply(all, function(data) { gam(nSteps ~ s(IntS, bs="cr") + Weekend, data=data, family=quasipoisson) })
lapply(fits, summary)
```
From the results, the result that there is a significant difference between weekends and weekdays remains invariant to which prediction is used.

## Weekday and Date

```{r weekday}
fitd <- gam(nSteps ~ s(IntS, bs="cr") + Date, data=data, family=quasipoisson)
summary(fitd)
fitw <- gam(nSteps ~ s(IntS, bs="cr") + Day + Date, data=data, family=quasipoisson)
summary(fitw)
```

* Interestingly, the Date coefficient is not significant. So this person seems rather consistent. 
* There are significative differences between Tuesday to Sunday vs Monday.

# Using q3()

Now I can proceed to summarize the results using q3() which basically fit the corresponding GAM model and makes a plot of the data.

```{r q3fun}
q3.weekend <- q3(data, "weekend")
summary(q3.weekend$fit)
q3.dow <- q3(data, "dow")
summary(q3.dow$fit)
```





# Reproducibility

```{r reproducibility}
sessionInfo()
print(proc.time())
```


